What We Eval in the Shadows

### 2021-04-14 paper/main.tex
emvirs

\noindent
As shown by Table~\ref{tab:highlevelenvironments}, most calls evaluate in a
function environment with the \c{global} kind as a distant second. This means
that most variable lookups and most side effects either read/update existing
local variables or introduce new ones. But for which function? From the point of
view of the function that called \eval, Table~\ref{tab:funoffset} gives an
offset on the call stack. Thus 0 is the direct caller and 1 is its parent, and
so on. The data suggests that in 81\% of cases, \eval access its caller's
environment -- this means, the variables of the function where \eval textually
occurs are read and written to. It is interesting that some 1.5\% of sites
evaluate a code in an environment that is three frame or above from \eval. This
distance implies that, in general, it is modular reasoning is impossible, to
understand the behavior of any piece of code entails fully understanding the
behavior of all functions that the code may call as the actions of \eval may
happen at a distance.

One site could appear in several categories. However, most sites, \ie
\packageNbOneCategoryEnvirSitePercent, have only kind of the above environments
passed to them, as shown in Table~\ref{tab:polyenvir}.

@@

The use of the global environment is likely split between intentional and
accidental use. Direct references using \c{globalenv()} or \c{.GlobalEnv} is
rare, only in \packageNbExplicitGlobalSites sites. We summarize that the
majority of uses of the global environment come from the fact that most of the
code in our corpus consists of scripts that are loaded at the top level, thus
any access of a caller frame may end up executing in the global environment. The
reason we make this point is that values that are stored in the global
environment are visible to all functions and are not reclaimed by the garbage
collector. So accidental uses could result in pollution of that name space.

Synthetic environments must be given a parent. When the \c{envir} argument is a list or a data frame, \eval also builds a new environment using the \c{enclos} argument as the parent. For \c{new.env}, the parent is passed explicitly at construction time. The parent is used to lookup
variables that are not found in the new environment (side effects go to the new
environment and not to its parent). Table~\ref{tab:newenvs} shows the parents of synthetic
environments. Over half are functions, the second most frequent case is packages.
??? Say more ????  FIX: so that we get to 100\%
\todo{It is over 100\% because one site can be passed different kinds of environments so it is a percentage over the number of sites in the dataset. I could also normalize to 100\% if it is more intuitive. In that case, the percentage would be over the sum of the rows (i.e. with duplicated sites).}

>>>>

\noindent
As shown by Table~\ref{tab:highlevelenvironments}, most calls evaluate in a
function environment with the \c{global} kind as a distant second. This means
that most variable lookups and most side effects either read/update existing
local variables or introduce new ones. But for which function? From the point of
view of the function that called \eval, Table~\ref{tab:funoffset} gives an
offset on the call stack. Thus 0 is the direct caller and 1 is its parent, and
so on. The data suggests that in 81\% of cases, \eval access its caller's
environment -- this means, the variables of the function where \eval textually
occurs are read and written to. It is interesting that some 1.5\% of sites
evaluate a code in an environment that is three frame or above from \eval. This
distance implies that, in general, it is modular reasoning is impossible, to
understand the behavior of any piece of code entails fully understanding the
behavior of all functions that the code may call as the actions of \eval may
happen at a distance. Finally note that any particular site may have several
kind of environments, but in \packageNbOneCategoryEnvirSitePercent of the case a
site has a single kind.

@@

\noindent
Sites that evaluate in the global environment are likely split between
intentional and accidental use. Direct references to the top-level, using
\c{globalenv()} or \c{.GlobalEnv}, are rare, they occur in only
\packageNbExplicitGlobalSites sites. We suspect that the majority of uses of the
global environment come from the fact that our corpus consists of scripts that
are loaded at the top level, thus that environment may often be the caller of
\eval or close to it. The reason we make this point is that values stored in the
global environment are visible to all functions and are not reclaimed by the
garbage collector. So accidental uses may pollute that name space.

Synthetic environments need a parent. For this, \eval uses its third argument
(\c{enclos}). When a synthetic environment is created by \eval, it has
\c{enclos} as a parent. The parent is used to lookup variables not found in its
child (side effects go to the child only). Table~\ref{tab:newenvs} shows the
kinds of the parents of synthetic environments. Over half are functions, then
come packages and global.
